---
title: "Example-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## How to Load the `beepr` Package

After installation, you can load the `beepr` package with the following command:

```{r setup}
library(beepr)
```

## Example Analysis

To demonstrate how to use the functions in the `beepr` package, we will conduct an example analysis.

### Question

How do U.S. Winter Olympic medalists' ages compare across sexes and sports?

### Original Data

The original data in this analysis is the [`olympics` dataset](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-07-27) from [TidyTuesday](https://www.tidytuesday.com).

#### Description

According to [the `olympics` dataset's documentation](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-07-27), the data contains information about athletic participants and medal wins in the modern Olympic Games (from Athens 1896 to Rio 2016).

#### Data Dictionary

The documentation for the `olympics` dataset provides [a helpful data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-07-27#data-dictionary).

### Analysis

#### Loading the Data into R

First, we need to load all necessary packages and the `olympics` dataset.

```{r packages, message = FALSE, warning = FALSE}
# Load the `here`, `tidytuesdayR`, and `tidyverse` packages
library(here)
library(tidytuesdayR)
library(tidyverse)

# Play a random sound once all of the packages have been loaded
beepr::beep(0)
```

```{r data, eval = FALSE}
# Test if a "data" directory exists locally; if not, create it
if(!dir.exists(here("data"))) { dir.create(here("data")) }

# Save the data only once (not each time the R Markdown document is knit)
if(!file.exists(here("data","olympics.rda"))) {
  url_csv <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv'
  olympics <- read_csv(url_csv)
  
  # Save the olympics file to an rda object
  save(olympics, file = here("data","olympics.rda"), version = 2)
}
```

```{r load, eval = FALSE, message = FALSE, warning = FALSE}
# Read in the data locally (each time the R Markdown document is knit)
olympics <- load(here::here("data","olympics.rda"))
```

#### Data Wrangling

To wrangle the `olympics` dataset, we will primarily use the `tidyverse` packages.

```{r wrangle, message = FALSE, warning = FALSE}
# Wrangling the Olympics data
wrangled_olympics <- olympics %>% 
  # Remove NA values from all columns (only leaves medal winners)
  tidyr::drop_na() %>% 
  # Filter to only include observations from US athletes, Winter games
  dplyr::filter(noc == "USA",
                season == "Winter") %>% 
  # Replace "M" with "Male", "F" with "Female", 
  dplyr::mutate(
    sex = dplyr::if_else(sex == "M", 
                         "Male", "Female")
  ) %>% 
  # Select only the variables relevant for this analysis
  dplyr::select(sex, sport, age) 
  
# Play a sound if glimpsing the new, wrangled dataset produces an error
beepr::beep_on_error(dplyr::glimpse(wrangled_olympics))
```

Before visualizing the data, we will explore it in a few different ways.

```{r purrr class, message = FALSE, warning = FALSE}
# Find the class of each column of the new, wrangled dataset
wrangled_olympics %>% 
  purrr::map_chr(class)
```
As expected `sex` and `sport` are character vectors, while `age` is a numeric vector. How many distinct observations are there in each vector?

```{r purrr distinct, message = FALSE, warning = FALSE}
# Find the number of distinct obs. in each column of the new, wrangled dataset
wrangled_olympics %>% 
  purrr::map_dbl(n_distinct)
```
It's interesting to note here that there are 13 different Winter Olympic sports represented in our dataset, which implies that there have been 13 different Winter Olympic sports that U.S. athletes have won a medal in.

#### Data Visualization

To visualize the `olympics` dataset, we will use the `ggplot2` package.

To start, let's look at how the average ages compare for male and female U.S. medalists at the Winter Olympic games.

```{r barplot, message = FALSE, warning = FALSE, fig.height=6, fig.width=8}
ggplot(data = wrangled_olympics, mapping = aes(x = sex, y = age)) +
  geom_bar(stat = "summary", fun.y = "mean", color = "black", fill = "black") +
  scale_y_continuous(label = scales::comma) +
  labs(title = "Spread of Mean U.S. Winter Olympic Medalist Ages by Sex",
       subtitle = "On average, male medalists tend to be slightly older than female medalists.",
       caption = "Data Source: TidyTuesday",
       x = "Sex",
       y = "Mean Age (Years)") +
  theme_minimal()
```

Although the means of male and female medalists' ages are similar, male medalists appear to be slightly older than female medalists.

To further answer our question, let's look at how U.S. medalists' ages compare by different winter sports.

```{r violinplot, message = FALSE, warning = FALSE, fig.height=6, fig.width=8}
ggplot(data = wrangled_olympics, mapping = aes(sport, age, fill = sport)) +
  geom_violin() +
  scale_y_continuous(label = scales::comma) +
  labs(title = "Spread of U.S. Winter Olympic Medalist Ages by Sport",
       subtitle = "Figure skating has the youngest winners, while ice hockey has the oldest.",
       caption = "Data Source: TidyTuesday",
       x = "Sport",
       y = "Distribution of Age (Years)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")
```

Regardless of sex, figure skating has the youngest winners, while ice hockey has the oldest among U.S. Winter Olympic medalists.

Now, let's look at both winter sports and U.S. medalist sexes. How are U.S. Winter Olympic medalists' ages spread by these two variables?

```{r facetplot, message = FALSE, warning = FALSE, fig.height=6, fig.width=8}
ggplot(data = wrangled_olympics, mapping = aes(sex, age, fill = sex)) +
  geom_boxplot(alpha = 0.5) +
  scale_y_continuous(label = scales::comma) +
  labs(title = "Spread of U.S. Winter Olympic Medalist Ages by Sex and Sport",
       subtitle = "In most sports, male winners tend to be older than female winners.",
       caption = "Data Source: TidyTuesday",
       x = "Sex",
       y = "Distribution of Age (Years)",
       fill = "Sex") +
  theme_minimal() +
  facet_wrap(vars(sport))
```

In most Winter Olympic sports, male U.S. medalists tend to be older than female U.S. medalists.
It's also interesting to note that some sports have not yet had any female medalists from the U.S. (e.g., Cross Country Skiing).

#### Summary of Results

In the majority of Winter Olympic sports, both the means and overall distributions of U.S. male medalists' ages are greater than those of U.S. female medalists' ages. 
Regardless of sex, figure skating as a winter sport has the youngest winners, while ice hockey has the oldest among all Winter Olympic U.S. medalists.

#### Functions Used

| **`dplyr`** | **`tidyr`** | **`purrr`** | **`ggplot2`** | **`beepr`** |
|------|------|------|------|------|
| filter() | drop_na() | map_chr() | geom_bar() | beep() |
| mutate() | | map_dbl() | geom_violin() | beep_on_error() |
| if_else() | | | geom_boxplot() | |
| select() | | | | |
| glimpse() | | | | |
